conmat: generate synthetic contact matrices for a given age-stratified population

### 2024-08-07 paper/paper.Rmd
adding some notes from review

There are currently few options for a user to generate their own synthetic contact matrices. 
Existing methods to generate synthetic contact matrices are not designed for replicability, do not have enough granularity, and does not cover enough administrative areas (in other words, some countries are not included).

[ What's different and useful about conmat? ] 
Users might have their own contact survey data that they would like to generate synthetic contact matrices from.
Perhaps the population demography is different, or the contact rates varying.
A higher level of granularity however is sometimes required to make public health decisions for a given population.

[ What else is covered in this paper? ] An example use-case for `conmat` is provided for a local government area (i.e. at the sub-national level) in Australia.
Also provided is an analysis pipeline to support conmat, [`syncomat`](https://github.com/idem-lab/syncomat), which generates synthetic contact matrices for 200 countries.

@@

#TODO - A better first sentence that encapsulates conmat use? 
Understanding the dynamics of infectious disease transmission is an important task (?) for epidemiologists and public policy makers.
Identifying vulnerable groups and predicting disease transmission (?)dynamics / how diseases spread are essential for informed public health decision-making. 
Infectious diseases such as influenza and coronavirus spread through human-to-human interactions, or in other words, "social contact". Quantifying social contact and its patterns can provide critical insights into how these diseases spread. [ Is this circular? ] / and how best to mitigate the spread of these diseases.

@@

These social contact surveys exist for a few countries. As an example, the "POLYMOD" study by @mossong2008 covered 8 European countries: Belgium, Germany, Finland, Great Britain, Italy, Luxembourg, The Netherlands, and Poland [@mossong2008]. However, what do we do when we want to estimate contact rates in other countries where this is not yet measured? We can use existing data--the contact rates obtained from contact surveys--to help us project / predict these estimates to countries or places that do not have them. These are called "synthetic contact matrices". A popular approach by @prem2017 projected contact rates from the POLYMOD study to 152 countries. This was later updated to include synthetic contact matrices for 177 countries at "urban" and "rural" levels for each country [@prem2021]. 
[ #TODO is project or predict a better word? Does it matter? ]

@@

The `conmat` package was developed to fill the specific need of creating synthetic contact matrices for local government areas in Australia. This package is used for [ #TODO what work, specifically? Health? Provide example. Or is *this* package commissioned by the Aus govt? ] work commissioned by the Australian government.
We developed methods and software to facilitate the following tasks.

@@

`conmat` was built to predict at four settings: work, school, home, and other. 
One model is fitted for each setting. 
Each model fitted is a Poisson generalised additive model (GAM) which predicts the count of contacts, with an offset for the log of participants. 
The model has six (?)covariates/terms to explain six key features of the relationship between ages, 
and two optional terms for attendance at school or work.
The two optional terms are included depending on which setting the model is fitted for.

Each cell in the resulting contact matrix, indexed *i*, *j*, is the predicted number of people in age group *j* that a single individual in age group *i* will have contact with per day. If you sum across all the *j* age groups for each *i* age group, you get the predicted total number of contacts per day for each individual of age group *i*. [ #TODO expected, predicted, or average? Does it matter? ]

The six terms are 
$|i-j|$, 
${|i-j|}^{2}$, 
$i + j$, 
$i \times j$, 
$\text{max}(i, j)$ and 
$\text{min}(i, j)$.

The six key features of the relationship between the age groups, represented by the six terms, are displayed in the figure below. 
[ #TODO notes-to-self: the model structure wasn't generated through any particularly robust process, it was just coming up with structures that looked mildly appropriate for our use case. ]

```{r}
# Show partial dep plot of the six main terms
```

Note that these partial dependency plots are on the log scale. 
When the six terms are added up together for each setting (in other words, each model) and exponentiated, they show the following patterns:

```{r}
# Show combined partial dep plot (i.e. sum of the partial dependencies for all six terms) in each setting: home, school, work and other
```

In other words, the six terms above provide patterns that are useful in modelling the different settings,
and correspond with real-life situations of how contact would look like. 
In the home setting for example, [ #TODO describe how children interact with parents and elderly generation, grandparents ].
When the terms for school and work are added, these terms also provide patterns that correspond with real-life situations.
https://idem-lab.github.io/conmat/dev/articles/visualising-conmat.html
In the school setting, children tend to contact other children in the same age groups as them.
In the work setting, there are no contacts with children under the age of ten and minimal contact with adults beyond retirement age.

@@

>>>>

There are currently few options for a user to generate their own synthetic contact matrices. [TODO add references/examples - e.g., socialmixr].
Existing methods to generate synthetic contact matrices are not designed for replicability, do not have enough granularity, and not all areas of interest (countries) are included.

[ What's different and useful about conmat? ]

Current interfaces to contact matrices are typically fixed to a given country. This means there aren't clear options for retrieving contact matrices for areas of different sizes. This is important as not all public health decisions are made at country levels - they might need to be made for specific smaller areas, such as cities and suburbs. Perhaps a smaller suburb within a country, or perhaps several countries all together. The conmat software allows users to provide their own age population data and combine this with information from an existing contact survey, such as POLYMOD (cite polymod). The software exposes model fitting and prediction separately to the user. This means users can generate synthetic contact matrices for any region they have age population information for.

[ What else is covered in this paper? ]

We demonstrate a use-case for `conmat` by creating contact matrices for a state (i.e. at the sub-national level) in Australia. It is also worth noting that there is an updated analysis pipeline that uses conmat, [`syncomat`](https://github.com/idem-lab/syncomat), to generate synthetic contact matrices for 200 countries.

@@

#TODO - A better first sentence that encapsulates conmat use?

Infectious diseases like influenza and COVID19 spread via social contact. If we can understand patterns of contact - which people are more likely be in contact with each other - then we will be able to create models of how disease spreads. Epidemiologists and public policy makers can uses these models to make decisions to keep a population safe and healthy.

Identifying vulnerable groups and predicting disease transmission (?)dynamics / how diseases spread are essential for informed public health decision-making. 
Infectious diseases such as influenza and coronavirus spread through human-to-human interactions, or in other words, "social contact". Quantifying social contact and its patterns can provide critical insights into how these diseases spread. [ Is this circular? ] / and how best to mitigate the spread of these diseases.

@@

These social contact surveys exist for a few countries. As an example, the "POLYMOD" study by @mossong2008 covered 8 European countries: Belgium, Germany, Finland, Great Britain, Italy, Luxembourg, The Netherlands, and Poland [@mossong2008]. However, what do we do when we want to estimate contact rates in other countries where this is not yet measured? We can use existing data--the contact rates obtained from contact surveys--to help us project / predict these estimates to countries or places that do not have them. These are called "synthetic contact matrices". A popular approach by @prem2017 projected contact rates from the POLYMOD study to 152 countries. This was later updated to include synthetic contact matrices for 177 countries at "urban" and "rural" levels for each country [@prem2021]. 
[ #TODO is project or predict a better word? Does it matter? ] (decision - use *predict*)

@@

The `conmat` package was developed to fill the specific need of creating synthetic contact matrices for local government areas in Australia. This package is used for [ #TODO what work, specifically? Health? Provide example. Or is *this* package commissioned by the Aus govt? note: @MikeLydeamore will reference the official government report] work commissioned by the Australian government.
We developed methods and software to facilitate the following tasks.

@@

`conmat` was built to predict at four settings: work, school, home, and other. 
One model is fitted for each setting. 
Each model fitted is a Poisson generalised additive model (GAM) which predicts the count of contacts, with an offset for the log of participants. 
The model has six covariates to explain six key features of the relationship between ages, 
and two optional covariates for attendance at school or work.
The two optional covariates are included depending on which setting the model is fitted for.

Each cell in the resulting contact matrix, indexed *i*, *j*, is the predicted number of people in age group *j* that a single individual in age group *i* will have contact with per day. If you sum across all the *j* age groups for each *i* age group, you get the predicted total number of contacts per day for each individual of age group *i*. [ #TODO expected, predicted, or average? Does it matter? note: currently going with predicted]

The six covariates are 
$|i-j|$, 
${|i-j|}^{2}$, 
$i + j$, 
$i \times j$, 
$\text{max}(i, j)$ and 
$\text{min}(i, j)$.

The six key features of the relationship between the age groups, represented by the six covariates, are displayed in the figure below. 
[ #TODO notes-to-self: the model structure wasn't generated through any particularly robust process, it was just coming up with structures that looked mildly appropriate for our use case. ]

```{r}
# Show partial dep plot of the six main covariates
```

Note that these partial dependency plots are on the log scale. 
When the six covariates are added up together for each setting (in other words, each model) and exponentiated, they show the following patterns:

```{r}
# Show combined partial dep plot (i.e. sum of the partial dependencies for all six covariates) in each setting: home, school, work and other
```

In other words, the six covariates above provide patterns that are useful in modelling the different settings,
and correspond with real-life situations of how contact would look like. 
In the home setting for example, [ #TODO describe how children interact with parents and elderly generation, grandparents ].
When the covariates for school and work are added, these covariates also provide patterns that correspond with real-life situations.
https://idem-lab.github.io/conmat/dev/articles/visualising-conmat.html
In the school setting, children tend to contact other children in the same age groups as them.
In the work setting, there are no contacts with children under the age of ten and minimal contact with adults beyond retirement age.

@@

# Cutting Room Floor
